---
title: "Introduction to Basic and Advanced Statistical Modelling"
subtitle: "From LMs to GLMMs"
author: "Alexandre Courtiol"
format:
   beamer:
       keep-tex: true
       aspectratio: 169
       toc: true
       classoption: t
fontsize: 10pt
knitr:
  opts_chunk: 
    dev: cairo_pdf
header-includes: |
  \setbeamertemplate{frametitle continuation}{}
  \setbeamertemplate{itemize items}[circle]
  \usepackage{caption}
  \usepackage{textpos}
  \setlength{\TPHorizModule}{1cm}
  \setlength{\TPVertModule}{1cm}
  
  % credit line
  \newcommand{\credit}{\scriptsize\textcolor{gray!70!black}{\quad courtiol@izw-berlin.de | \url{www.datazoogang.de} | \url{www.izw-berlin.de/en}}}
  
  % create footer with slide number
  \setbeamertemplate{footline}{%
    \credit\hfill\usebeamercolor[fg]{page number in head/foot}%
    \usebeamerfont{page number in head/foot}%
    \insertframenumber{} / \inserttotalframenumber\kern1em\vskip2pt%
  }
  
  \usepackage{colortbl}
  \usepackage{makecell}
  \usepackage{tcolorbox}
  \tcbuselibrary{skins}
  \usepackage{etoolbox}
  \usepackage{forest}
  \usepackage{amssymb}
  
  % make the red boxes around important stuff
  \newtcolorbox{callout}{%
    colback=white, 
    colframe=red, 
    boxrule=1pt, 
    arc=6pt, 
    boxsep=2pt, 
    left=8pt, 
    right=8pt, 
    top=4pt, 
    bottom=4pt,
    before skip=6pt, % space before the box
    after skip=6pt   % space after the box
  }
  
  % 95% CI
  \newcommand{\ci}{\ensuremath{\text{CI}_{\text{95\%}}}}
  
  % highlights
  \newcommand{\bluenorm}[1]{\textcolor{bluebeamer}{#1}}
  \newcommand{\blueit}[1]{\textcolor{bluebeamer}{\emph{#1}}}
  \newcommand{\bluebold}[1]{\textcolor{bluebeamer}{\textbf{#1}}}
  
  % define some custom colours
  \definecolor{okgreen}{HTML}{00CC99}
  \definecolor{bad}{HTML}{FF5050}
  \definecolor{bluebeamer}{HTML}{3333B3}
  \usepackage{xcolor} % for other colours
  
  \captionsetup{labelformat=empty, justification=centering}
  
  % for displaying section slides (with center alignment)
  \AtBeginSection[]{ 
    \begin{frame}[c]
    \centering
    \Large\insertsectionhead
    \end{frame}
  }
  
  % for grey background behind in-line R code
  \let\oldtexttt\texttt
  \renewcommand{\texttt}[1]{\colorbox{gray!20}{\oldtexttt{#1}}}
editor_options:
  chunk_output_type: console
---

```{r}
#| label: setup
## load packages
start <- Sys.time() # for recording duration
library(tidyverse)
library(brms) # explicit call here since can be hidden by cache below

## set tibble sizes via pillar
options(width = 90, pillar.width = 90,
        pillar.print_min = 5, pillar.print_max = 5)

## set output text size
knitr::opts_chunk$set(size = "footnotesize")
def.chunk.hook <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  if (!is.null(options$size) && options$size != "normalsize") {
    paste0("\n\\", options$size, "\n\n", x, "\n\n\\normalsize")
  } else {
    x
  }
})

## set base size in ggplot2
theme_set(theme_classic(base_size = 30))

## counter for practice slides, to be called with `r next_practice()`
practice_counter <- 0
next_practice <- function(counter = NULL) {
  if (!is.null(counter)) 
    practice_counter <<- counter
  else
    practice_counter <<- practice_counter + 1
  practice_counter
}

## create cache folder for bmrs
if (!dir.exists("cache_brms")) dir.create("cache_brms")
```

# Linear Models (LM)

## LM --- what are models?

\vfill

\bluebold{What is a statistical model?}

a \blueit{statistical model} is an approximate description of the \blueit{data generating process}

. . .

\vfill

\bluebold{What is a linear model?}

a data generating process produced by a \blueit{linear function}: the data are generated from a set of terms by multiplying each term by a constant (a model parameter) and summing them

. . .

\vfill

**NB:** seeing models as data generating process is both conceptually and practically useful

- it helps for understanding what model parameters mean
- it helps for estimating statistical power
- it helps for knowing how to test fitting procedure by simulation

\vfill

## LM --- notations of a LM: simple notation

\vspace{1em}

\begin{center}
$y_i = \beta_0 + \beta_1 \times x_{1,i} + \beta_2 \times x_{2,i} + \dots + \beta_{p} \times x_{p,i} + \epsilon_i$
\end{center}

. . .

- $y_i$ = the response variable / dependent variable / observations to explain
- $\beta_j$ = the model parameters / regression coefficients
- $x_{j,i}$ = regressors / constants derived from the predictor variables (aka predictors for short)
- $\epsilon_i$ = the errors / residual errors

. . .

\vspace{1em}

**NB:**

- this is the notation that biologists should use in their paper/thesis to describe their model
- the regressors are sometimes directly given by the columns from your dataset (i.e. the predictors), but not always (more on that later)


## LM --- notations of a LM: usual R notation

\vspace{1em}

`y ~ x_1 + x_2 + ... + x_p`

\vspace{1em}

**NB:**

- the parameters and the error term are not written down


## LM --- notations of a LM: matrix notation

\vspace{1em}

\begin{center}
$Y = X \beta + \epsilon$
\end{center}

$$
\begin{bmatrix} y_1 \\ y_2 \\ y_3 \\ \dots \\ y_n \end{bmatrix} =
\begin{bmatrix}
1 & x_{1,1} & x_{1,2} & \dots & x_{1,p} \\
1 & x_{2,1} & x_{2,2} & \dots & x_{2,p} \\
1 & x_{3,1} & x_{3,2} & \dots & x_{3,p} \\
\dots & \dots & \dots & \dots & \dots \\
1 & x_{n,1} & x_{n,2} & \dots & x_{n,p}
\end{bmatrix}
\begin{bmatrix}
\beta_0 \\ \beta_1 \\ \beta_2 \\ \dots \\ \beta_p
\end{bmatrix} +
\begin{bmatrix}
\epsilon_1 \\ \epsilon_2 \\ \epsilon_3 \\ \dots \\ \epsilon_n
\end{bmatrix}
$$

. . .

- $Y$ = the vector of observations
- $X$ = a matrix called the design (or model) matrix
- $\beta$ = the vector of model parameters
- $\epsilon$ = the vector of errors

\vspace{1em}

**NB:** this is the notation for statisticians and computer scientists, which is sometimes used to code Bayesian models


## LM --- data generating process
<!-- Emerald Lion is an anagram to linear model -->

Let's enter \bluebold{God(dess) mode} $\rightarrow$ we pretend we know how the world generates data

. . .

\bluenorm{Emerald Lion} (a goddess) informed us that she grows aliens of a certain size as follows:

```{=latex}
\begin{equation*}
\begin{cases}
\text{size}_i = 50 + 1.5 \times \text{humans\_eaten}_{i} + \epsilon_i \\
\epsilon_i \sim \mathcal{N}(0, \sigma^2 = 4)
\end{cases}
\end{equation*}
```

. . .

We can thus generate the size of 10 aliens that have eaten humans following E.L.'s recipe\
(we assume that the number of humans eaten follows a Poisson with $\lambda = 3$):

```{r}
#| label: alien data
#| echo: true
set.seed(123L)
N <- 10
Alien <- tibble(humans_eaten = rpois(N, lambda = 3), # any alternative would do 
                intercept = 50, slope = 1.5,
                error = rnorm(n = N, mean = 0, sd = 2), # must be rnorm!
                size = intercept + slope*humans_eaten + error)
```

. . . 

**NB:** we could also write the last 2 lines as:

```{r}
#| label: alien data2
#| echo: true
#| eval: false
size <- rnorm(n = 10, mean = intercept + slope*humans_eaten, sd = 2)
```

## LM --- data generating process

\bluebold{Simulated data}
```{r}
#| label: alien study
#| echo: true
Alien
```

\textcolor{red}{can you recognise:}

- the \blueit{model parameters}?
- the \blueit{regressor} (here = predictor)?
- the \blueit{errors}?
- the \blueit{response variable}?


## LM --- model fitting

In practice, model parameters are unknown, so we have to estimate them\
(that is the whole point of statistical modelling...)

- in frequentist statistics, \blueit{fitting the model} means estimating the values of the model parameters that maximise the probability of the data

\vspace{1em}

```{r}
#| label: alien lm fake
#| echo: true
#| eval: false
fit_alien_lm <- lm(size ~ humans_eaten, data = Alien)
summary(fit_alien_lm)
```

```{r}
#| label: alien lm real
#| echo: false
#| size: tiny
fit_alien_lm <- lm(size ~ humans_eaten, data = Alien)
summary(fit_alien_lm)
```

## LM --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

Let's first use {brms}, which is simpler than {nimble} but handles a smaller set of models:
```{r}
#| label: alien lm brms fake
#| echo: true
#| eval: false
#| message: false
library(brms) # Bayesian regression models using stan
fit_alien_lm_brms <- brm(size ~ humans_eaten, data = Alien)
summary(fit_alien_lm_brms)
```

```{=html}
<!~~
if compilation errors occur on the slide (e.g. cmath: No such file or directory),
try running the block below before rendering,
delete the knitr cache and re-render
~~>
```

```{r}
#| label: alien lm brms real
#| echo: false
#| message: false
#| cache: true
#| size: tiny
library(brms) # Bayesian regression models using stan
fit_alien_lm_brms <- brm(size ~ humans_eaten, data = Alien,
                         seed = 123,
                         file = "cache_brms/fit_alien_lm_bmrs") # to store the compiled model (quicker to rerun)
summary(fit_alien_lm_brms)
```

## LM --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

Let's now use {nimble}:
```{r}
#| label: alien lm nimble prep
#| echo: true
#| size: scriptsize
library(nimble)
library(posterior)
library(bayesplot)

code <- nimbleCode({ 
  # model mimicking the data generating process
  for (i in seq_len(N)) {
      size[i] ~ dnorm(mean = intercept + slope*humans_eaten[i], sd = sigma)
  }
  # prior
  intercept ~ dnorm(mean = 50, sd = 20)
  slope ~ dnorm(mean = 0, sd = 10)
  sigma ~ dunif(min = 0, max = 100)
})

constants <- list(N = nrow(Alien),
                  humans_eaten = Alien$humans_eaten)

inits <- list(intercept = mean(Alien$size), slope = 0, sigma = 1)
```

## LM --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

Before fitting the model, let's make sure our priors for the intercept and slope are not too vague

::: columns
::: {.column width="60%"}
```{r}
#| echo: true
#| eval: false
set.seed(123)
n_sims <- 100
prior_intercept <- rnorm(n_sims, mean = 50, sd = 20) 
prior_slope <- rnorm(n_sims, mean = 0, sd = 10)
humans_eaten_possible <-
  min(Alien$humans_eaten):max(Alien$humans_eaten)

expand_grid(sim = seq_len(n_sims),
            humans_eaten = humans_eaten_possible) |>
  mutate(size = prior_intercept[sim] +
           prior_slope[sim] * humans_eaten) -> prior_pred

ggplot(prior_pred, aes(humans_eaten, size)) +
  geom_line(aes(group = sim), linewidth = 0.1) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(group = NULL), data = Alien)
```
:::
::: {.column width="40%"}
```{r}
#| echo: false
#| out-width: 0.9\linewidth
#| fig-align: center
set.seed(123)
n_sims <- 100
prior_intercept <- rnorm(n_sims, mean = 50, sd = 20) 
prior_slope <- rnorm(n_sims, mean = 0, sd = 10)
humans_eaten_possible <-
  min(Alien$humans_eaten):max(Alien$humans_eaten)

expand_grid(sim = seq_len(n_sims),
            humans_eaten = humans_eaten_possible) |>
  mutate(size = prior_intercept[sim] +
           prior_slope[sim] * humans_eaten) -> prior_pred

ggplot(prior_pred, aes(humans_eaten, size)) +
  geom_line(aes(group = sim), linewidth = 0.1) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(group = NULL), data = Alien)
```

. . .

- this looks good: not too many silly sizes and yet wide enough to cover a wide range of slopes
- scaling the data would improve things (more on that soon) 
:::
:::


## LM --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

\vspace{1em}

```{r}
#| label: alien lm nimble fit fake
#| echo: true
#| eval: false
fit_alien_lm_nimb <- nimbleMCMC(code = code,
                                constants = constants,
                                data = list(size = Alien$size), # response goes here
                                inits = inits)

fit_alien_lm_nimb_d <- as_draws(fit_alien_lm_nimb) # to clean and check for real use
summary(fit_alien_lm_nimb_d)
```

```{r}
#| label: alien lm nimble fit real
#| echo: false
#| cache: true
#| size: tiny
fit_alien_lm_nimb <- nimbleMCMC(code = code,
                                constants = constants,
                                data = list(size = Alien$size),
                                inits = inits,
                                nchains = 1,
                                setSeed = 123,
                                progressBar = FALSE)

fit_alien_lm_nimb_d <- as_draws(fit_alien_lm_nimb)
summary(fit_alien_lm_nimb_d)
```

**NB:** 

- in real use, the predictor variables should be scaled (more on that soon) & the posteriors should be trimmed, thinned and checked


## LM --- model fitting results

Let's compare the results

::: columns
::: {.column width="50%"}
`lm()`
```{r}
#| echo: false
#| label: alien lm results
#| size: scriptsize
rbind(cbind(summary(fit_alien_lm)$coefficients,
            confint(fit_alien_lm))[, -c(3:4)],
      sigma = c(summary(fit_alien_lm)$sigma,
                sqrt(fit_alien_lm$df.residual/2 * (2*summary(fit_alien_lm)$sigma^2/fit_alien_lm$df.residual)^2), #shape * scale^2
                qgamma(p = c(0.025, 0.975),
                       shape = fit_alien_lm$df.residual/2,
                       scale = 2*summary(fit_alien_lm)$sigma^2/fit_alien_lm$df.residual))) |> 
  as.data.frame() |> 
  rownames_to_column(var = "Parameter") |> 
  as_tibble()
``` 
`brms::brm()`
```{r}
#| echo: false
#| label: alien lm brms results
#| size: scriptsize
rbind(fixef(fit_alien_lm_brms),
      sigma = as.data.frame(VarCorr(fit_alien_lm_brms)[[1]][[1]])) |> 
  as.data.frame() |> 
  rownames_to_column(var = "Parameter") |> 
  as_tibble()
```
`nimble::nimbleMCMC()`
```{r}
#| echo: false
#| label: alien lm nimble results
#| size: scriptsize
summarise_draws(fit_alien_lm_nimb_d, "mean", "sd", ~ quantile(.x, probs = c(0.025, 0.975)))[c(1, 3, 2), ]
```
:::

. . .

::: {.column width="50%"}
::: {.incremental}
- we never obtained $50$ (intercept), $1.5$ (slope), or $2$ (sigma) because a sample is unlikely to reflect perfectly the population-level values; only N=$\infty$ or talking to the goddess could get you those
- despite large uncertainties (`{r} paste("N=", N)`) the methods are rather consistent and not too far off
- in `lm()` $\hat{\sigma}$ is usually not a parameter that is estimated directly, it is deduced once the intercept and slope(s) have been fitted (behind the scenes I struggled to compute the SE and CI for it)
- by default {nimble} returns q5 and q95 (not q2.5 and q97.5), so I had to request this manually (see `?summarise_draws`)
:::
:::
:::

## Practice `r next_practice()`


# Questions? {.unnumbered}

## {.c}
\centering
\Large Questions?


# Developer corner {.unnumbered}

## {.c}
\centering
\Large Developer corner


## Ignore this slide --- (technical info for Alex)

```{r}
#| echo: false
#| size: "tiny"
options(width = 130)
duration_min <- round(as.numeric(difftime(Sys.time(), start, units = "min")), digits = 1)
paste(duration_min, "min of R running time")
print(sessionInfo(), locale = FALSE, tzone = FALSE)
```