---
title: "Introduction to Basic and Advanced Statistical Modelling"
subtitle: "From LMs to GLMMs"
author: "Alexandre Courtiol"
format:
   beamer:
       keep-tex: true
       aspectratio: 169
       toc: true
       classoption: t
fontsize: 10pt
knitr:
  opts_chunk: 
    dev: cairo_pdf
header-includes: |
  \setbeamertemplate{frametitle continuation}{}
  \setbeamertemplate{itemize items}[circle]
  \usepackage{caption}
  \usepackage{textpos}
  \setlength{\TPHorizModule}{1cm}
  \setlength{\TPVertModule}{1cm}
  
  % credit line
  \newcommand{\credit}{\scriptsize\textcolor{gray!70!black}{\quad courtiol@izw-berlin.de | \url{www.datazoogang.de} | \url{www.izw-berlin.de/en}}}
  
  % create footer with slide number
  \setbeamertemplate{footline}{%
    \credit\hfill\usebeamercolor[fg]{page number in head/foot}%
    \usebeamerfont{page number in head/foot}%
    \insertframenumber{} / \inserttotalframenumber\kern1em\vskip2pt%
  }
  
  \usepackage{colortbl}
  \usepackage{makecell}
  \usepackage{tcolorbox}
  \tcbuselibrary{skins}
  \usepackage{etoolbox}
  \usepackage{forest}
  \usepackage{amssymb}
  
  % make the red boxes around important stuff
  \newtcolorbox{callout}{%
    colback=white, 
    colframe=red, 
    boxrule=1pt, 
    arc=6pt, 
    boxsep=2pt, 
    left=8pt, 
    right=8pt, 
    top=4pt, 
    bottom=4pt,
    before skip=6pt, % space before the box
    after skip=6pt   % space after the box
  }
  
  % 95% CI
  \newcommand{\ci}{\ensuremath{\text{CI}_{\text{95\%}}}}
  
  % highlights
  \newcommand{\bluenorm}[1]{\textcolor{bluebeamer}{#1}}
  \newcommand{\blueit}[1]{\textcolor{bluebeamer}{\emph{#1}}}
  \newcommand{\bluebold}[1]{\textcolor{bluebeamer}{\textbf{#1}}}
  
  % define some custom colours
  \definecolor{okgreen}{HTML}{00CC99}
  \definecolor{bad}{HTML}{FF5050}
  \definecolor{bluebeamer}{HTML}{3333B3}
  \usepackage{xcolor} % for other colours
  
  \captionsetup{labelformat=empty, justification=centering}
  
  % for displaying section slides (with center alignment)
  \AtBeginSection[]{ 
    \begin{frame}[c]
    \centering
    \Large\insertsectionhead
    \end{frame}
  }
  
  % for grey background behind in-line R code
  \let\oldtexttt\texttt
  \renewcommand{\texttt}[1]{\colorbox{gray!20}{\oldtexttt{#1}}}
editor_options:
  chunk_output_type: console
---

```{r}
#| label: setup
## load packages
start <- Sys.time() # for recording duration
library(tidyverse)
library(brms) # explicit call here since can be hidden by cache below

## set tibble sizes via pillar
options(width = 90, pillar.width = 90,
        pillar.print_min = 5, pillar.print_max = 5)

## set output text size
knitr::opts_chunk$set(size = "footnotesize")
def.chunk.hook <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  if (!is.null(options$size) && options$size != "normalsize") {
    paste0("\n\\", options$size, "\n\n", x, "\n\n\\normalsize")
  } else {
    x
  }
})

## set base size in ggplot2
theme_set(theme_classic(base_size = 30))

## counter for practice slides, to be called with `r next_practice()`
practice_counter <- 0
next_practice <- function(counter = NULL) {
  if (!is.null(counter)) 
    practice_counter <<- counter
  else
    practice_counter <<- practice_counter + 1
  practice_counter
}

## create cache folder for bmrs
if (!dir.exists("cache_brms")) dir.create("cache_brms")
```

# Linear Models introduction (LM)

## LM intro --- what are models?

\vfill

\bluebold{What is a statistical model?}

a \blueit{statistical model} is an approximate description of the \blueit{data generating process}

. . .

\vfill

\bluebold{What is a linear model?}

a data generating process produced by a \blueit{linear function}: the data are generated from a set of terms by multiplying each term by a constant (a model parameter) and summing them

. . .

\vfill

**NB:** seeing models as data generating process is both conceptually and practically useful

- it helps for understanding what model parameters mean
- it helps for estimating statistical power
- it helps for knowing how to test fitting procedure by simulation

\vfill

## LM intro --- notations of a LM: simple notation

\vspace{1em}

\begin{center}
$y_i = \beta_0 + \beta_1 \times x_{1,i} + \beta_2 \times x_{2,i} + \dots + \beta_{p} \times x_{p,i} + \epsilon_i$
\end{center}

. . .

- $y_i$ = the response variable / dependent variable / observations to explain
- $\beta_j$ = the model parameters / regression coefficients
- $x_{j,i}$ = regressors / constants derived from the predictor variables (aka predictors for short)
- $\epsilon_i$ = the errors / residual errors

. . .

\vspace{1em}

**NB:**

- this is the notation that biologists should use in their paper/thesis to describe their model
- the regressors are sometimes directly given by the columns from your dataset (i.e. the predictors), but not always (more on that later)


## LM intro --- notations of a LM: usual R notation

\vspace{1em}

`y ~ x_1 + x_2 + ... + x_p`

\vspace{1em}

**NB:**

- the parameters and the error term are not written down


## LM intro --- notations of a LM: matrix notation

\vspace{1em}

\begin{center}
$Y = X \beta + \epsilon$
\end{center}

$$
\begin{bmatrix} y_1 \\ y_2 \\ y_3 \\ \dots \\ y_n \end{bmatrix} =
\begin{bmatrix}
1 & x_{1,1} & x_{1,2} & \dots & x_{1,p} \\
1 & x_{2,1} & x_{2,2} & \dots & x_{2,p} \\
1 & x_{3,1} & x_{3,2} & \dots & x_{3,p} \\
\dots & \dots & \dots & \dots & \dots \\
1 & x_{n,1} & x_{n,2} & \dots & x_{n,p}
\end{bmatrix}
\begin{bmatrix}
\beta_0 \\ \beta_1 \\ \beta_2 \\ \dots \\ \beta_p
\end{bmatrix} +
\begin{bmatrix}
\epsilon_1 \\ \epsilon_2 \\ \epsilon_3 \\ \dots \\ \epsilon_n
\end{bmatrix}
$$

. . .

- $Y$ = the vector of observations
- $X$ = a matrix called the design (or model) matrix
- $\beta$ = the vector of model parameters
- $\epsilon$ = the vector of errors

\vspace{1em}

**NB:** this is the notation for statisticians and computer scientists, which is sometimes used to code Bayesian models


## LM intro --- data generating process
<!-- Emerald Lion is an anagram to linear model -->

Let's enter \bluebold{God(dess) mode} $\rightarrow$ we pretend we know how the world generates data

. . .

\bluenorm{Emerald Lion} (a goddess) informed us that she grows aliens of a certain size as follows:

```{=latex}
\begin{equation*}
\begin{cases}
\text{size}_i = 50 + 1.5 \times \text{humans\_eaten}_{i} + \epsilon_i \\
\epsilon_i \sim \mathcal{N}(0, \sigma^2 = 4)
\end{cases}
\end{equation*}
```

. . .

We can thus generate the size of 10 aliens that have eaten humans following E.L.'s recipe\
(we assume that the number of humans eaten follows a Poisson with $\lambda = 3$):

```{r}
#| label: alien data
#| echo: true
set.seed(123L)
N <- 10
Alien <- tibble(humans_eaten = rpois(N, lambda = 3), # any alternative would do 
                intercept = 50, slope = 1.5,
                error = rnorm(n = N, mean = 0, sd = 2), # must be rnorm!
                size = intercept + slope*humans_eaten + error)
```

. . . 

**NB:** we could also write the last 2 lines as:

```{r}
#| label: alien data2
#| echo: true
#| eval: false
size <- rnorm(n = 10, mean = intercept + slope*humans_eaten, sd = 2)
```

## LM intro --- data generating process

\bluebold{Simulated data}
```{r}
#| label: alien study
#| echo: true
Alien
```

- which columns show the \blueit{model parameters}?
- which columns show \blueit{regressor} (here = predictor)?
- which columns show \blueit{errors}?
- which columns show \blueit{response variable}?
- how the \blueit{response variable} is computed?
- what would the \blueit{mean response value} be, for an alien who does not eat humans?
- what would the \blueit{mean response value} be, for an alien who eats 2 humans?


## LM intro --- model fitting

In practice, model parameters are unknown, so we have to estimate them\
(that is the whole point of statistical modelling...)

- in frequentist statistics, \blueit{fitting the model} means estimating the values of the model parameters that maximise the probability of the data

\vspace{1em}

```{r}
#| label: alien lm fake
#| echo: true
#| eval: false
fit_alien_lm <- lm(size ~ humans_eaten, data = Alien)
summary(fit_alien_lm)
```

```{r}
#| label: alien lm real
#| echo: false
#| size: tiny
fit_alien_lm <- lm(size ~ humans_eaten, data = Alien)
summary(fit_alien_lm)
```

## LM intro --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

Let's first use {brms}, which is simpler than {nimble} but handles a smaller set of models:
```{r}
#| label: alien lm brms fake
#| echo: true
#| eval: false
#| message: false
library(brms) # Bayesian regression models using stan
fit_alien_lm_brms <- brm(size ~ humans_eaten, data = Alien)
summary(fit_alien_lm_brms)
```

```{=html}
<!~~
if compilation errors occur on the slide (e.g. cmath: No such file or directory),
try running the block below before rendering,
delete the knitr cache and re-render
~~>
```

```{r}
#| label: alien lm brms real
#| echo: false
#| message: false
#| cache: true
#| size: tiny
library(brms) # Bayesian regression models using stan
fit_alien_lm_brms <- brm(size ~ humans_eaten, data = Alien,
                         seed = 123,
                         file = "cache_brms/fit_alien_lm_bmrs") # to store the compiled model (quicker to rerun)
summary(fit_alien_lm_brms)
```

## LM intro --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

Let's now use {nimble}:
```{r}
#| label: alien lm nimble prep
#| echo: true
#| size: scriptsize
library(nimble)
library(posterior)
library(bayesplot)

code <- nimbleCode({ 
  # model mimicking the data generating process
  for (i in seq_len(N)) {
      size[i] ~ dnorm(mean = intercept + slope*humans_eaten[i], sd = sigma)
  }
  # prior
  intercept ~ dnorm(mean = 50, sd = 20)
  slope ~ dnorm(mean = 0, sd = 10)
  sigma ~ dunif(min = 0, max = 100)
})

constants <- list(N = nrow(Alien),
                  humans_eaten = Alien$humans_eaten)

inits <- list(intercept = mean(Alien$size), slope = 0, sigma = 1)
```

## LM intro --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

Before fitting the model, let's make sure our priors for the intercept and slope are not too vague

::: columns
::: {.column width="65%"}
```{r}
#| echo: true
#| eval: false
set.seed(123)
n_sims <- 100
prior_intercept <- rnorm(n_sims, mean = 50, sd = 20) 
prior_slope <- rnorm(n_sims, mean = 0, sd = 10)
humans_eaten_possible <-
  min(Alien$humans_eaten):max(Alien$humans_eaten)

expand_grid(sim = seq_len(n_sims),
            humans_eaten = humans_eaten_possible) |>
  mutate(size = prior_intercept[sim] +
           prior_slope[sim] * humans_eaten) -> prior_pred

ggplot(prior_pred, aes(humans_eaten, size)) +
  geom_line(aes(group = sim), linewidth = 0.1) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(group = NULL), data = Alien)
```
:::
::: {.column width="35%"}
```{r}
#| echo: false
#| out-width: 0.9\linewidth
#| fig-align: center
set.seed(123)
n_sims <- 100
prior_intercept <- rnorm(n_sims, mean = 50, sd = 20) 
prior_slope <- rnorm(n_sims, mean = 0, sd = 10)
humans_eaten_possible <-
  min(Alien$humans_eaten):max(Alien$humans_eaten)

expand_grid(sim = seq_len(n_sims),
            humans_eaten = humans_eaten_possible) |>
  mutate(size = prior_intercept[sim] +
           prior_slope[sim] * humans_eaten) -> prior_pred

ggplot(prior_pred, aes(humans_eaten, size)) +
  geom_line(aes(group = sim), linewidth = 0.1) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(group = NULL), data = Alien)
```

. . .

- this looks good: not too many silly sizes and yet wide enough to cover a wide range of slopes
- scaling the data would improve things (more on that soon) 
:::
:::


## LM intro --- model fitting

- in Bayesian statistics, \blueit{fitting the model} means using the data to characterise the posterior distribution

\vspace{1em}

```{r}
#| label: alien lm nimble fit fake
#| echo: true
#| eval: false
fit_alien_lm_nimb <- nimbleMCMC(code = code,
                                constants = constants,
                                data = list(size = Alien$size), # response goes here
                                inits = inits)

fit_alien_lm_nimb_d <- as_draws(fit_alien_lm_nimb) # to clean and check for real use
summary(fit_alien_lm_nimb_d)
```

```{r}
#| label: alien lm nimble fit real
#| echo: false
#| cache: true
#| size: tiny
fit_alien_lm_nimb <- nimbleMCMC(code = code,
                                constants = constants,
                                data = list(size = Alien$size),
                                inits = inits,
                                nchains = 1,
                                setSeed = 123,
                                progressBar = FALSE)

fit_alien_lm_nimb_d <- as_draws(fit_alien_lm_nimb)
summary(fit_alien_lm_nimb_d)
```

**NB:** 

- in real use, the predictor variables should be scaled (more on that soon) & the posteriors should be trimmed, thinned and checked


## LM intro --- model fitting results

Let's compare the results

::: columns
::: {.column width="50%"}
`lm()`
```{r}
#| echo: false
#| label: alien lm results
#| size: scriptsize
rbind(cbind(summary(fit_alien_lm)$coefficients,
            confint(fit_alien_lm))[, -c(3:4)],
      sigma = c(summary(fit_alien_lm)$sigma,
                sqrt(fit_alien_lm$df.residual/2 * (2*summary(fit_alien_lm)$sigma^2/fit_alien_lm$df.residual)^2), #shape * scale^2
                qgamma(p = c(0.025, 0.975),
                       shape = fit_alien_lm$df.residual/2,
                       scale = 2*summary(fit_alien_lm)$sigma^2/fit_alien_lm$df.residual))) |> 
  as.data.frame() |> 
  rownames_to_column(var = "Parameter") |> 
  as_tibble()
``` 
`brms::brm()`
```{r}
#| echo: false
#| label: alien lm brms results
#| size: scriptsize
rbind(fixef(fit_alien_lm_brms),
      sigma = as.data.frame(VarCorr(fit_alien_lm_brms)[[1]][[1]])) |> 
  as.data.frame() |> 
  rownames_to_column(var = "Parameter") |> 
  as_tibble()
```
`nimble::nimbleMCMC()`
```{r}
#| echo: false
#| label: alien lm nimble results
#| size: scriptsize
summarise_draws(fit_alien_lm_nimb_d, "mean", "sd", ~ quantile(.x, probs = c(0.025, 0.975)))[c(1, 3, 2), ]
```
:::

. . .

::: {.column width="50%"}
::: {.incremental}
- we never obtained $50$ (intercept), $1.5$ (slope), or $2$ (sigma) because a sample is unlikely to reflect perfectly the population-level values; only N=$\infty$ or talking to the goddess could get you those
- despite large uncertainties (`{r} paste("N=", N)`) the methods are rather consistent and not too far off
- in `lm()` $\hat{\sigma}$ is usually not a parameter that is estimated directly, it is deduced once the intercept and slope(s) have been fitted (behind the scenes I struggled to compute the SE and CI for it)
- by default {nimble} returns q5 and q95 (not q2.5 and q97.5), so I had to request this manually (see `?summarise_draws`)
:::
:::
:::

## Practice `r next_practice()`

- using the `penguins` dataset (readily available in R), fit a linear model with:
  - `body_mass` as the response variables
  - `flipper_len` as predictors

. . . 

\vfill

**NB:** do not forget to clean and plot the dataset

```{r}
#| echo: true
#| label: penguins2
#| out-width: 4cm
#| fig-align: center
penguins |> 
  drop_na(body_mass, flipper_len) -> penguins2

ggplot(penguins2) +
  geom_point(aes(y = body_mass, x = flipper_len))
```

# Interpreting coefficients

## Interpreting coefficients --- intercept and slopes

\textcolor{red}{How to interpret the regression coefficients?}

```{r}
#| echo: true
#| label: fit flipper
fit_flipper <- lm(body_mass ~ flipper_len, data = penguins2)
coef(fit_flipper)
```

. . .

- the \blueit{intercept} is the predicted response when all regressors are at 0
```{r}
#| echo: true
#| label: fit flipper interpretation int
predict(fit_flipper, newdata = list(flipper_len = 0))
```

## Interpreting coefficients --- intercept and slopes

\textcolor{red}{How to interpret the regression coefficients?}

```{r}
#| echo: true
#| label: fit flipper bis
fit_flipper <- lm(body_mass ~ flipper_len, data = penguins2)
coef(fit_flipper)
```

- a \blueit{slope} gives the change in predicted response when its associated regressor increases by 1 unit (here 1 mm)
```{r}
#| echo: true
#| label: fit flipper interpretation slope
pred_diff <- predict(fit_flipper, newdata = list(flipper_len = c(200, 201)))
pred_diff
pred_diff[2] - pred_diff[1] # or simply diff(pred_diff)
```


## Interpreting coefficients --- scaling predictors

\textcolor{red}{How to interpret the regression coefficients when predictors are scaled?}

\blueit{scaling} means removing the mean (i.e. \blueit{centering}) and divide by the SD (i.e. \blueit{standardising}), i.e. it means transforming the variable into \blueit{z-scores}
```{r}
#| echo: true
#| label: prep flipper scaled
penguins2 |> 
  mutate(flipper_len_z = scale(flipper_len)) -> penguins3
```

**NB:** 

- scaling both the response variable and the predictors is a recommended practice when using {nimble}, {stan}, etc... since it makes the definition of priors more consistent across studies; it also improves the efficiency of some MCMC samplers considerably since it decreases correlations between posteriors
- some packages do similar things internally (e.g. {brms}), while doing themselves the necessary \blueit{back transformation} for allowing interpretations at the original scales



## Interpreting coefficients --- scaling predictors

\textcolor{red}{How to interpret the regression coefficients when predictors are scaled?}
```{r}
#| echo: true
#| label: fit flipper scaled
fit_flipper_z <- lm(body_mass ~ flipper_len_z, data = penguins3)
coef(fit_flipper_z)
```

. . .

- the \blueit{intercept} remains the predicted response when all regressors are at 0, but `flipper_len = 0` meant a flipper length of 0 mm, while `flipper_len_z = 0` means an average flipper length (i.e. `mean(penguins3$flipper_len)` = `{r} round(mean(penguins3$flipper_len), 1)` mm)


## Interpreting coefficients --- scaling predictors

\textcolor{red}{How to interpret the regression coefficients when predictors are scaled?}
```{r}
#| echo: true
#| label: fit flipper scaled 2
fit_flipper_z <- lm(body_mass ~ flipper_len_z, data = penguins3)
coef(fit_flipper_z)
```

- a \blueit{slope} remains the change in predicted response when its associated regressor increases by 1 unit, but here 1 unit is not longer 1 mm but 1 SD in flipper length (i.e. `sd(penguins3$flipper_len)` = `{r} round(sd(penguins3$flipper_len), 1)` mm)

\bluebold{back transformation}

```{r}
#| echo: true
#| label: equivalence flip
coef(fit_flipper_z)["flipper_len_z"] / sd(penguins3$flipper_len)
```


## Interpreting coefficients --- scaling the response

\textcolor{red}{How to interpret the regression coefficients when the response is scaled?}

```{r}
#| echo: true
#| label: prep body mass scaled
penguins3 |> 
  mutate(body_mass_z = scale(body_mass)) -> penguins4
```

```{r}
#| echo: true
#| label: fit body mass scaled
fit_flipper_bmz <- lm(body_mass_z ~ flipper_len, data = penguins4)
coef(fit_flipper_bmz)
```

- again, the \blueit{intercept} remains the predicted response when all regressors are at 0, and a \blueit{slope} remains the change in predicted response when its associated regressor increases by 1 unit, but this time the response is expressed in SDs above the mean (i.e. z-scores) rather than grams

## Interpreting coefficients --- scaling the response

\textcolor{red}{How to interpret the regression coefficients when the response is scaled?}

```{r}
#| echo: true
#| label: prep body mass scaled 2
penguins3 |> 
  mutate(body_mass_z = scale(body_mass)) -> penguins4
```

```{r}
#| echo: true
#| label: fit body mass scaled 2
fit_flipper_bmz <- lm(body_mass_z ~ flipper_len, data = penguins4)
coef(fit_flipper_bmz)
```

\bluebold{back transformation}

```{r}
#| echo: true
#| label: equivalence flip bm
coef(fit_flipper_bmz)["(Intercept)"] * sd(penguins4$body_mass) + mean(penguins4$body_mass)

coef(fit_flipper_bmz)["flipper_len"] * sd(penguins4$body_mass)
```

## Practice `r next_practice()`

\textcolor{red}{How to interpret the regression coefficients when the response and predictors are scaled?}

```{r}
#| echo: true
#| label: fit all scaled
fit_flipper_zz <- lm(body_mass_z ~ flipper_len_z, data = penguins4)
coef(fit_flipper_zz)
```

1. interpret the slope on its new scale
2. interpret the intercept on its new scale
3. apply the required back transformation to perform the interpretation of the slope on the original scale
4. think about how you would define the priors in this case


## Solution

\textcolor{red}{How to interpret the regression coefficients when the response and predictors are scaled?}

```{r}
#| echo: true
#| label: fit all scaled 2
fit_flipper_zz <- lm(body_mass_z ~ flipper_len_z, data = penguins4)
coef(fit_flipper_zz)
```

1. the slope describes the increase in SD of body mass for an increase of 1 SD in flipper length
2. the intercept is the mean scaled value of the body mass for a penguin with average flipper length, which is why it is 0
3. back transformation of the slope:

```{r}
#| echo: true
#| label: fit all scaled solution
(coef(fit_flipper_zz)["flipper_len_z"] / sd(penguins4$flipper_len)) * sd(penguins4$body_mass)
```

## Solution

4. $\mathcal{N}(0, \sigma = 2)$ for both the intercept and slope seems appropriate

::: columns
::: {.column width="65%"}
```{r}
#| label: prior test z fake
#| echo: true
#| eval: false
set.seed(123)
n_sims <- 100
prior_intercept <- rnorm(n_sims, mean = 0, sd = 2) 
prior_slope <- rnorm(n_sims, mean = 0, sd = 2)
flipper_z <- -3:3

expand_grid(sim = seq_len(n_sims),
            flipper_len_z = flipper_z) |>
  mutate(body_mass_z = prior_intercept[sim] +
           prior_slope[sim] * flipper_len_z) -> prior_pred

ggplot(prior_pred, aes(flipper_len_z, body_mass_z)) +
  geom_line(aes(group = sim), linewidth = 0.1) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(group = NULL), data = penguins4)
```
:::
::: {.column width="35%"}
```{r}
#| label: prior test z real
#| echo: false
#| eval: true
#| out-width: 0.9\linewidth
#| fig-align: center
set.seed(123)
n_sims <- 100
prior_intercept <- rnorm(n_sims, mean = 0, sd = 2) 
prior_slope <- rnorm(n_sims, mean = 0, sd = 2)
flipper_z <- -3:3

expand_grid(sim = seq_len(n_sims),
            flipper_len_z = flipper_z) |>
  mutate(body_mass_z = prior_intercept[sim] +
           prior_slope[sim] * flipper_len_z) -> prior_pred

ggplot(prior_pred, aes(flipper_len_z, body_mass_z)) +
  geom_line(aes(group = sim), linewidth = 0.1) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(group = NULL), data = penguins4)
```
:::
:::



## Interpreting coefficients --- qualitative predictors

\textcolor{red}{How to interpret the regression coefficients?}

```{r}
#| echo: true
#| label: fit sex
fit_sex <- lm(body_mass ~ sex, data = penguins)
coef(fit_sex)
```



## co plots

```{r}
#| label: coplot intro
#| echo: true
#| out-width: 3cm
#| fig-align: center
coplot(body_mass ~ flipper_len | species + sex, panel = panel.smooth, data = penguins2)
```

# Questions? {.unnumbered}

## {.c}
\centering
\Large Questions?


# Developer corner {.unnumbered}

## {.c}
\centering
\Large Developer corner


## Ignore this slide --- (technical info for Alex)

```{r}
#| echo: false
#| size: "tiny"
options(width = 130)
duration_min <- round(as.numeric(difftime(Sys.time(), start, units = "min")), digits = 1)
paste(duration_min, "min of R running time")
print(sessionInfo(), locale = FALSE, tzone = FALSE)
```