---
title: "Introduction to Basic and Advanced Statistical Modelling"
subtitle: "Using R (solutions to exercises)"
author: "Alexandre Courtiol"
format:
   beamer:
       keep-tex: true
       aspectratio: 169
       toc: true
       toc-depth: 2
       classoption: t
header-includes: |
  \setbeamertemplate{frametitle continuation}{}
  \setbeamertemplate{itemize items}[circle]
  \usepackage{caption}
  \captionsetup{labelformat=empty, justification=centering}
  \usepackage{textpos}
  \setlength{\TPHorizModule}{1cm}
  \setlength{\TPVertModule}{1cm}
  \setbeamertemplate{footline}{%
    \hfill\usebeamercolor[fg]{page number in head/foot}%
    \usebeamerfont{page number in head/foot}%
    \insertframenumber{} / \inserttotalframenumber\kern1em\vskip2pt%
  }
  \usepackage{xcolor}
  \let\oldtexttt\texttt
  \renewcommand{\texttt}[1]{\colorbox{gray!20}{\oldtexttt{#1}}}
  
  \AtBeginSection[]{ % for centering the section slides since others are top aligned
    \begin{frame}[c]
    \centering
    \Large\insertsectionhead
    \end{frame}
  }
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
## set tibble sizes via pillar
options(width = 90, pillar.width = 90,
        pillar.print_min = 4, pillar.print_max = 4)

## set output text size
knitr::opts_chunk$set(size = "footnotesize")
def.chunk.hook <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  if (!is.null(options$size) && options$size != "normalsize") {
    paste0("\n\\", options$size, "\n\n", x, "\n\n\\normalsize")
  } else {
    x
  }
})

## counter for practice slides, to be called with `r next_practice()`
practice_counter <- 0
next_practice <- function(counter = NULL) {
  if (!is.null(counter)) 
      practice_counter <<- counter
  else
    practice_counter <<- practice_counter + 1
  practice_counter
}

library(tidyverse)
library(sf)
species_count <- read_csv("data_MHB/header-species-count.csv")
species_count_sf <- st_as_sf(species_count, coords = c("lon", "lat"), crs = 4326)
```

# Reading data

## Practice `r next_practice(4)`
```{r}
#| echo: true
#| message: true
territories <- read_csv("data_MHB/data-territories.csv")
```

# Wrangling data

## Practice `r next_practice(5)`a

Use `select()` and/or `filter()` and/or `slice()` and/or `arrange()` to determine:

1. the largest number of recorded species across all years

```{r}
#| eval: false
#| echo: true
#| message: true
species_count |> 
    arrange(desc(n_recorded_species)) |> 
    slice(1) |> 
    select(n_recorded_species)
```

or

```{r}
#| eval: false
#| echo: true
#| message: true
species_count |> 
    filter(n_recorded_species == max(n_recorded_species)) |> 
    select(n_recorded_species)
```

## Practice `r next_practice(5)`b

Use `select()`, `filter()`, `slice()` and/or `arrange()` to determine:

2. the largest number of recorded species in 2025


```{r}
#| eval: false
#| echo: true
#| message: true
species_count |> 
    filter(year == 2025) |> 
    filter(n_recorded_species == max(n_recorded_species)) |> 
    select(n_recorded_species)
```

but NOT this:

```{r}
#| eval: false
#| echo: true
#| message: true
species_count |> 
    filter(year == 2025, n_recorded_species == max(n_recorded_species)) |>
    select(n_recorded_species)
```
Do you understand why?
<!-- if both are used together, it is like an AND and the condition is here never satisfied -->  


## Practice `r next_practice(6)`a

Use the functions we have seen to obtain:

1. the range of median elevation according to the number of visits

```{r}
#| echo: true
#| message: true
species_count |> 
    summarise(median_elevation_min = min(median_elevation),
              median_elevation_max = max(median_elevation),
              .by = n_visits)
```


## Practice `r next_practice(6)`b

Use the functions we have seen to obtain:

2. the 3 most recorded species for each year

```{r}
#| echo: true
#| message: true
territories |> 
    summarise(territories_sum = sum(territories_total),
              .by = c(year, nameeg, namelt)) |> # aggregate across samplearea_id
    slice_max(territories_sum, n = 3, by = year) |> # pick top 3 per year
    select(year, nameeg, namelt, territories_sum)
```


## Practice `r next_practice(6)`c (1)

Use the functions we have seen to obtain:

3. the number of observations for *Sturnus vulgaris* for each year

```{r}
#| echo: true
#| message: true
territories |> 
    filter(namelt == "Sturnus vulgaris") |> 
    summarise(territories_sum = sum(territories_total, na.rm = TRUE), # NAs present
              .by = year)
```

## Practice `r next_practice(6)`c (2)

Use the functions we have seen to obtain:

3. the number of observations for *Sturnus vulgaris* for each year

```{r}
#| echo: true
#| message: true
territories |> 
    filter(namelt == "Sturnus vulgaris") |> 
    drop_na(territories_total) |> # alternative not to use na.rm below
    summarise(territories_sum = sum(territories_total), # NAs present
              .by = year)
```

## Practice `r next_practice(7)`

Try to use `pivot_longer()` to get back to the original long format of the data from the output of the following code:

```{r}
#| echo: true
#| message: true
territories |> 
    filter(year == 2025) |> 
    select(samplearea_id, nameeg, territories_total) |> 
    pivot_wider(names_from = nameeg, 
                values_from = territories_total) |> 
    pivot_longer(-samplearea_id,
                 names_to = "nameeg",
                 values_to = "territories_total")
```


# Wrangling spatial data

## Practice `r next_practice(8)`a

Use the functions we have seen to obtain:

1. the sampled area (sensu convex hull) across years

```{r}
#| echo: true
#| message: true
species_count_sf |> 
    summarise(geometry = st_combine(geometry), .by = year) |> 
    mutate(hull = st_convex_hull(geometry),
           area = st_area(hull),
           area_km2 = units::set_units(area, "km^2")) |> 
    select(-hull) |> 
    st_drop_geometry()
```


## Practice `r next_practice(8)`b (1)

Use the functions we have seen to obtain:

2. the IDs of the sampled areas which are most peripheral

Simple sorting by distance to centroid (but lacks threshold)
```{r}
#| eval: true
#| echo: true
#| message: true
species_count_sf |> 
  distinct(samplearea_id, geometry) |> 
  mutate(centroid = st_centroid(st_combine(geometry)),
         dist = st_distance(centroid, geometry, by_element = TRUE)) |> 
  arrange(desc(dist)) |> 
  as_tibble() # (for slide only -> skip metadata clutter)
```


## Practice `r next_practice(8)`b (2)

More useful, but more complex: intersection with concave hull outline
```{r}
#| echo: true
#| message: true
species_count_sf |> 
  distinct(samplearea_id, .keep_all = TRUE) |> 
  mutate(all_points = st_combine(geometry), # combine points into new geometry
         hull = st_concave_hull(all_points, ratio = 0.2), # create concave hull
         perimeter = st_boundary(hull), # extract outer line of the hull
         perimeter = st_buffer(perimeter, dist = 1000), # create buffer zone around it
         on_perim = st_intersects(perimeter, geometry, # test if points fall into buffer
                                  by_element = TRUE)) -> data_perimeter_sf 
  
data_perimeter_sf |> 
    filter(on_perim) |> 
    pull(samplearea_id) # pull() extracts a column
```


## Practice `r next_practice(8)`b (2)

More useful, but more complex: intersection with concave hull outline
```{r}
#| echo: true
#| message: true
#| out-height: "50%"
#| fig-align: "center"
theme_set(theme_bw(base_size = 20)) # setting the theme for all plots
ggplot(data_perimeter_sf) +
  geom_sf(aes(geometry = perimeter)) +
  geom_sf(aes(colour = on_perim)) # geometry = geometry is implicit
```


# Plotting data

## Practice `r next_practice(10)` (1)

- combine data wrangling and plotting to illustrate the change in species abundance across years for the top 6 most encountered species

```{r}
#| echo: true
#| message: true
territories |> 
    summarise(territories_sum = sum(territories_total, na.rm = TRUE),
              .by = c(nameeg, namelt)) |> 
    slice_max(territories_sum, n = 6) |> 
    pull(namelt) -> top6sp

top6sp
```

## Practice `r next_practice(10)` (2)

- combine data wrangling and plotting to illustrate the change in species abundance across years for the top 6 most encountered species

```{r}
#| echo: true
#| message: true
territories |> 
    filter(namelt %in% top6sp) |> 
    summarise(territories_sum = sum(territories_total, na.rm = TRUE),
              .by = c(nameeg, namelt, year)) -> top6_across_time

top6_across_time
```

## Practice `r next_practice(10)` (3a)

- combine data wrangling and plotting to illustrate the change in species abundance across years for the top 6 most encountered species

```{r}
#| echo: true
#| message: true
#| fig-width: 15
#| out-width: "85%"
#| fig-align: "center"
ggplot(top6_across_time) +
  geom_line(aes(y = territories_sum, x = year, colour = nameeg))
```


## Practice `r next_practice(10)` (3b)

- combine data wrangling and plotting to illustrate the change in species abundance across years for the top 6 most encountered species

```{r}
#| echo: true
#| message: true
#| fig-width: 15
#| out-height: "60%"
#| fig-align: "center"
ggplot(top6_across_time) +
  geom_line(aes(y = territories_sum, x = year)) +
  facet_wrap(~ nameeg, scales = "free_y")
```
